### 2017-01-30 manuscript.rmd
Update manuscript.rmd
Minor edits

Given the ortholog conjecture, we would expect to see larger contrasts following duplication events than speciation events. Instead, we find no such distribution (Figure XXKMRR). The Wilcox rank test does not reject the null hypothesis that the rate of evolution following duplications is the same as or less than the rate following speciation (p value = `r wilcox_test_result`). Our results therefore find no evidence for the ortholog conjecture in this system.

@@

In order to better understand why our phylogenetic analysis seems to support a different conclusion (*i.e.*, no support for the ortholog conjecture) than the published analysis of KMRR [@KryuchkovaMostacci:2016iw] (*i.e.*, strong support for the ortholog conjecture), we first checked to make sure we could reproduce their result based on pairwise analyses. This is important since we are only looking at at a subset of the data they considered, the Brawand *et al.* 2011 [@Brawand:2011du] dataset for gene trees that could be successfully time calibrated. We can replicate their result - we find a higher Tau correlation between orthologs (R=`r round( ortholog_r, 3) `) than between paralogs (R=`r round( paralog_r, 3)`).

Why is it that pairwise methods and phylogenetic methods lead to opposite conclusions? The reason is that they describe different things. Pairwise comparisons describe contemporary patterns, while phylogenetic methods infer historical processes. There need not be a different process of evolution following speciation and duplication for paralogs to be more different than orthologs. This could be due to the structure of the gene phylogenies. If paralogs tend to be more distantly related to each other than orthologs, then there would be more time for differences to accumulate even if the rate of change is the same between the two. This is, in fact, the case for these data. While the mean distance between orthologs is `r round( ortholog_distance_mean, 1 )` My, the mean distance between paralogs is `r round( paralog_distance_mean, 1 )` My. As described above, this is because the oldest speciation event is by definition the most recent common ancestor of the the species included in the study, but many gene families underwent duplication before this time.

@@

The ortholog conjecture does not have to be an all or nothing question. It may be the case that the rates of phenotypic evolution following duplication may be greater than that following duplication in some organisms, gene families, and evolutionary processes. We just don't find evidence for it when summarizing gene expression across tissues with Tau in these mammals. This calls into question the general predictive power of the ortholog conjecture, and until these processes are better understood it will be necessary to test for it in each situation. These tests should be articulated in terms of clear alternative hypotheses [@Studer:2009bm] that make distinct phylogenetic comparative predictions.

@@

We are not the first to apply phylogenetic comparative methods to functional genomic data. While the vast majority of comparative functional genomics studies have used pairwise similarity methods, a small number of CFG studies have employed phylogenetic comparative approaches [@Oakley:2005ky; @Eng:2009gp; @Chang:2014hq]. For instance, a phylogenetic ANOVA [@Rohlfs:2015bd] of the evolution of gene expression improves statistical power and drastically reduces the rate of false positives relative to pairwise approaches.

>>>>

Given the ortholog conjecture, we would expect to see larger contrasts following duplication events than speciation events. Instead, we find no such distribution (Figure XXKMRR). The Wilcoxon rank test does not reject the null hypothesis that the rate of evolution following duplications is the same as or less than the rate following speciation (p value = `r wilcox_test_result`). Our results therefore find no evidence for the ortholog conjecture in this system.

@@

In order to better understand why our phylogenetic analysis supports a different conclusion (*i.e.*, no support for the ortholog conjecture) than the published analysis of KMRR [@KryuchkovaMostacci:2016iw] (*i.e.*, strong support for the ortholog conjecture), we first checked to make sure we could reproduce their result based on pairwise analyses. This is important since we are only looking at at a subset of the data they considered, the Brawand *et al.* 2011 [@Brawand:2011du] dataset for gene trees that could be successfully time calibrated. We can replicate their result - we find a higher Tau correlation between orthologs (R=`r round( ortholog_r, 3) `) than between paralogs (R=`r round( paralog_r, 3)`).

Why is it that pairwise methods and phylogenetic methods lead to opposite conclusions? The reason is that they describe different things. Pairwise comparisons describe contemporary patterns, while phylogenetic methods infer historical processes [@OMeara:2012gq]. There need not be a different process of evolution following speciation and duplication for paralogs to be more different than orthologs. This could be due to the structure of the gene phylogenies. If paralogs tend to be more distantly related to each other than orthologs, then there would be more time for differences to accumulate even if the rate of change is the same between the two. This is, in fact, the case for these data. While the mean distance between orthologs is `r round( ortholog_distance_mean, 1 )` My, the mean distance between paralogs is `r round( paralog_distance_mean, 1 )` My. As described above, this is because the oldest speciation event is by definition the most recent common ancestor of the species included in the study, but many gene families underwent duplication before this time.

@@

The ortholog conjecture does not have to be an all or nothing question. It may be the case that the rates of phenotypic evolution following duplication may be greater than that following duplication in some organisms, gene families, and evolutionary processes [@Gabaldon:2013fi]. We just don't find evidence for it when summarizing gene expression across tissues with Tau in these mammals. This calls into question the general predictive power of the ortholog conjecture, and until these processes are better understood it will be necessary to test for it in each situation. These tests should be articulated in terms of clear alternative hypotheses [@Studer:2009bm] that make distinct phylogenetic comparative predictions.

@@

We are not the first to apply phylogenetic comparative methods to functional genomic data. While the vast majority of comparative functional genomics studies have used standard pairwise similarity methods, a small number of CFG studies have employed phylogenetic comparative approaches [@Oakley:2005ky; @Eng:2009gp; @Chang:2014hq]. For instance, a phylogenetic ANOVA [@Rohlfs:2015bd] of the evolution of gene expression improves statistical power and drastically reduces the rate of false positives relative to pairwise approaches.
